---
title: "CIM"
date: "23 January 2017"
output: html_document
---
```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(dpi = 100, echo= TRUE, warning=FALSE, message=FALSE, #dev = 'jpeg',
                      fig.show=TRUE, fig.keep = 'all', fig.height= 8, fig.width=9)
```


# Clustered Image Maps

…is based on a hierarchical clustering simultaneously applied on the rows and the columns of a real-valued similarity matrix. This similarity matrix is obtained via the results of rCCA or (s)PLS. The matrix is graphically represented as a 2-dimensional colored image, where each entry of the matrix is colored on the basis of its value, and where the rows and columns are reordered according to a hierarchical clustering. Dendrograms resulting from the clustering are added to the left (or right) side and to the top (or bottom) of the image.

# Usage in mixOmics

CIM can be obtained in **mixOmics** via the function *cim()* as displayed below:
```{r}
library(mixOmics)
```

```{r}
set.seed(123)
    mat <- matrix(rnorm(200), 20, 10)
    mat[ 1:10, seq(1, 10, 2)] <- mat[ 1:10, seq(1, 10, 2)] - 4
    mat[11:20, seq(2, 10, 2)] <- mat[11:20, seq(2, 10, 2)] + 2
    mat[15:20, seq(2, 10, 2)] <- mat[15:20, seq(2, 10, 2)] + 3
    colnames(mat) <- paste("row", 1:10, sep = " ")
    rownames(mat) <- paste("gene", 1:20, sep = "")
#test col.sideColors
    cond=rep(c("cond 1","cond 2"),5)
    cond.col <- c("cond 1" = "darkviolet", "cond 2" = "darkorange")
    obj.cim=cim(mat)
#test row.sideColors
    ddr <- as.hclust(obj.cim$ddr)
    cl <- cutree(ddr, k = 3)
    gene.col <- c("up" = "red", "down" = "green", "null" = "black")

    cim(mat, col.sideColors = cond.col[cond],row.sideColors = gene.col[cl],
    legend=list( legend = names(gene.col), 
    col = gene.col,title = "Condition", cex = 0.7))
```

```{r}
data(vac18.simulated)
    X <- vac18.simulated$genes
    Y <- vac18.simulated$stimulation
    design <- data.frame(samp = vac18.simulated$sample,
    stim = vac18.simulated$stimulation,
    time = vac18.simulated$time)
    
    res.2level <- splsda(X, Y, ncomp = 2, keepX = c(120, 10))
    
    stim.col <- c("darkblue", "purple", "green4","red3")
    stim.col <- stim.col[as.numeric(design$stim)]
    time.col <- c("orange", "cyan")[as.numeric(design$time)]
    
    cim(res.2level, row.sideColors = cbind(stim.col, time.col),
    row.names = paste(design$time, design$stim, sep = "_"),
    col.names = FALSE,
    legend=list(title="Stimulations"))
```

```{r}
data(liver.toxicity)

repeat.indiv <- c(1, 2, 1, 2, 1, 2, 1, 2, 3, 3, 4, 3, 4, 3, 4, 4, 5, 6, 5, 5,
    6, 5, 6, 7, 7, 8, 6, 7, 8, 7, 8, 8, 9, 10, 9, 10, 11, 9, 9,
    10, 11, 12, 12, 10, 11, 12, 11, 12, 13, 14, 13, 14, 13, 14,
    13, 14, 15, 16, 15, 16, 15, 16, 15, 16)
    
design <- data.frame(sample = vac18$sample)
    
res.spls.1level <- spls(X = liver.toxicity$gene,
    Y=liver.toxicity$clinic,
    ncomp = 3,
    keepX = c(50, 50, 50),
    keepY = c(5,5,5),
    mode = "canonical")
    
col <- c("darkblue", "purple", "green4","red3")
cim(res.spls.1level,
    mapping="Y",
    row.names=liver.toxicity$treatment[,3],
    row.sideColors = col[factor(liver.toxicity$treatment[,3])],
    legend=list(legend=unique(liver.toxicity$treatment[,3]),
    col=col))
```

# References 

[González I., Lê Cao K.-A., Davis, M.D. and Déjean S. (2013) Insightful graphical outputs to explore relationships between two ‘omics’ data sets. BioData Mining 5:19](http://mixomics.org/wp-content/uploads/2012/03/Visualising_JDataMining.pdf)

    