---
title: "Missing Values"
date: "20 January 2017"
output: html_document
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(dpi = 100, echo= TRUE, warning=FALSE, message=FALSE, #dev = 'jpeg',
                      fig.show=TRUE, fig.keep = 'all')
```

# Missing values

All methodologies implemented in **mixOmics** can handle missing values. In particular, **(s)PLS**, **(s)PLS-DA**, **(s)PCA** (using the non-linear iterative partial least squares algorithm **NIPALS** approach) take advantage of the PLS algorithm which performs local regressions on the latent components (NIPALS, Wold 1966).

The **valid()** function for *(s)PLS-DA* has not yet been implemented to deal with missing values. A solution for now is to impute the missing values from each data set separately using the **nipals()** function.

# Usage in mixOmics

```{r}
library(mixOmics)
data(liver.toxicity)
X <- liver.toxicity$gene[, 1:100] # a reduced size data set

## pretend there are 20 NA values in our data
na.row <- sample(1:nrow(X), 20, replace = TRUE)
na.col <- sample(1:ncol(X), 20, replace = TRUE)
X.na <- as.matrix(X)

## fill these NA values in X
X.na[cbind(na.row, na.col)] <- NA
sum(is.na(X.na)) # Should display 20

# this might take some time depending on the size of the data set
nipals.tune = nipals(X.na, reconst = TRUE, ncomp = 10)$eig
barplot(nipals.tune, xlab = 'number of components', ylab = 'explained variance')

#nipals with the chosen number of components (try to choose a large number)
nipals.X = nipals(X.na, reconst = TRUE, ncomp = 10)$rec

#  only replace the imputation for the missing values
id.na = is.na(X.na)
nipals.X[!id.na] = X[!id.na]

nipals.X[id.na]  # imputed values
X[id.na]         # original values
```

# References 

Wold H. (1966) Multivariate Analysis. Academic Press, New York, Wiley.

Tenenhaus M. (1998) La régression PLS : théorie et pratique. Editions Technip.