---
title: "Networks"
date: "23 January 2017"
output: html_document
---
```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(dpi = 100, echo= TRUE, warning=FALSE, message=FALSE, #dev = 'jpeg',
                      fig.show=TRUE, fig.keep = 'all', fig.height= 8, fig.width=9)
```

# Networks

Networks can be displayed to represent the relevant associations between X and Y variables, they are inferred using a pairwise association score between X and Y variables. The user can set a threshold to represent only the variables X and Y with an association score greater than the threshold. By changing this threshold, the user can choose to include or exclude relationships in the relevance network. The output is a graph where each X- and Y-variable corresponds to a node and the edges included in the graph portray associations between them.

# Usage in mixOmics

Networks graphs can be obtained in mixOmics via the function network as displayed below:
```{r}
library(mixOmics)
```

```{r}
## Will work with sPLS result
data(liver.toxicity)
    
    X <- liver.toxicity$gene
    Y <- liver.toxicity$clinic
    toxicity.spls <- spls(X, Y, ncomp = 3)
                          
 network(toxicity.spls,
         color.node = c("orange","blue"),
         cutoff = 0.75,
         row.names = FALSE, 
         col.names = FALSE)
```
 
```{r}
network(toxicity.spls,color.edge = c("red","orange","cyan","blue"),row.names=FALSE, cutoff = 0.75)
```
 
```{r}
 network(toxicity.spls,lty.edge='dotted',row.names = FALSE, cutoff = 0.75)
```

We can export the graph to **Cytoscape** file format using write.graph from the igraph package. 
 
```{r eval = FALSE}
net.result <- network(toxicity.spls)

write.graph(net.result$gR, file = "network.gml", format = "gml")
```

# References 

[González I., Lê Cao K.-A., Davis, M.D. and Déjean S. (2013) Insightful graphical outputs to explore relationships between two ‘omics’ data sets. BioData Mining 5:19](http://mixomics.org/wp-content/uploads/2012/03/Visualising_JDataMining.pdf)
 
 