---
title: "plotArrow"
date: "23 January 2017"
output: html_document
---
```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(dpi = 100, echo= TRUE, warning=FALSE, message=FALSE, #dev = 'jpeg',
                      fig.show=TRUE, fig.keep = 'all', fig.height= 8, fig.width=9)
```

# Arrow plot

Each arrow corresponds one sample.  The start of the arrow indicates the location of the sample in X in one plot, and the tip the location of the sample in Y in the other plot. Short arrows indicate if both data sets strongly agree and long arrows a disagreement between the two data sets. the latent variables (or ‘variates’) are given as an input from a **(s)pls object**.

# Usage in mixOmics

```{r}
library(mixOmics)
```


```{r}
#SGCCA

diet = unmap(nutrimouse$diet)
blocks = list(gene = nutrimouse$gene, 
              lipid = nutrimouse$lipid, 
              diet = diet)

design = matrix(c(0,1,1,1,0,1,1,1,0), ncol = 3, nrow = 3, byrow = TRUE)
    
nutri.sgcca <- wrapper.sgcca(blocks,design=design, ncomp = 3)

#test blocks with ind.names vector and group
plotArrow(nutri.sgcca, 
          ind.names= nutrimouse$diet, 
          group=nutrimouse$diet,
          legend=TRUE)
```

```{r}
#PLS/SPLS

data(liver.toxicity)
    
X <- liver.toxicity$gene
Y <- liver.toxicity$clinic
toxicity.spls <- spls(X, Y, ncomp = 3,
keepX = c(50, 50, 50), keepY = c(10, 10, 10))
    

###position start , group and abline
    
plotArrow(toxicity.spls,
          ind.names= liver.toxicity$treatment[, 3],
          position.names='start',
          group= liver.toxicity$treatment[, 4],
          abline = TRUE)
```

```{r}
#RCC
data(nutrimouse)
    
X <- nutrimouse$lipid
Y <- nutrimouse$gene
nutri.res <- rcc(X, Y, 
                 ncomp = 3, 
                 lambda1 = 0.064, 
                 lambda2 = 10)

plotArrow(nutri.res,
          col = color.mixo(as.numeric(nutrimouse$genotype)),
          pch=c(0,1),
          cex=c(1,2))
```

