---
title: "WPBloggingFromR to mixOmics site"
author: "Danielle Davenport"
date: "23 November 2016"
output: html_document
---
Steps to post your RMarkdown file to WordPress Mixomics Blog  

These are the required Libraries and Packages
#Step 1: Check libraries and packages 
Might need: 
install.packages("devtools")
install.packages("RCurl")
install.packages("XML")
devtools:::install_github("duncantl/XMLRPC")
devtools:::install_github("duncantl/RWordPress")
```{r}
library(RWordPress)
library(knitr)
# Set figure dimensions
#opts_chunk$set(fig.width=5, fig.height=5)
# Set figures to upload to imgur.com
opts_knit$set(upload.fun = imgur_upload, base.url = NULL) 
```


(a) The Easy Way
If you are using Rstudio make an Rmd document and then click *knit*

#Step 3
Now setup wordpress details
```{r}
# Set your WP username, password, and your site URL
options(WordpressLogin = c(your.username = 'your.password'), #update these
        WordpressURL = 'https://mixomics.org/xmlrpc.ph')
```
#Step 4
Post a new blog post 
```{r}
knit2wp(input = 'your Rmd File.Rmd', title = 'The title of your post',shortcode = FALSE ,publish = FALSE)

#publish = FALSE this keeps the post as a draft so you can check it on wordpress before publishing 
```

#Step 5
Now check post on mixomics page
*it will be posted as a post, not a page. To change it to a page, just go to the mixomics site, find the post in the "Posts" tab, click "quick edit"", and change the "post type" dropdown box from post -> page. you can now find it under the "page" tab.
Then go to the pages tab, review, then publish. 
Now add to menu in "Apperence/Menu"

-------
References/Links
------

http://3.14a.ch/archives/2015/03/08/how-to-publish-with-r-markdown-in-wordpress/

http://francojc.github.io/publishing-rmarkdown-to-wordpress-or-jekyll/

http://chinapleth.io/2016/01/13/write-posts-with-rstudio-rmarkdown-format-and-publish-directly-to-wordpress-with-knitr-rwordpress/